MSDS 6306: Doing Data Science - Case Study 01 - Beers/Breweries Dataset

### 2020-10-13 MSDS6306_DDS_CaseStudy01-Beersdataset.Rmd
RMD

---
title: "MSDS 6306: Doing Data Science - Case Study 01 - Beers/Breweries Dataset"
author: "Renfeng Wang & Venkata 'MG' Vanga"
date: "9/30/2020"
output: html_document
  chunk_output_type: console
---

@@

```{r libraries}
library(dplyr)
library(ggplot2)
library(tidyr)
require(graphics)
library(tidyverse)
library(magrittr)
library(GGally)
library(ggthemes)
library(plotly)
library(maps)
library(stringr)
library(stringi)
library(mapproj)
library(RCurl)
library(readr)
library(rio)

#Beer and Breweries data input from CSV files
```{r Beer Data}
# This reads in the Beers data from select folder file Beers.csv.
Beerdata <- read.csv("https://raw.githubusercontent.com/VenkataVanga/MSDS-6306/main/Beers.csv",header = T,sep = ",",na.strings = "NA",fill = TRUE);
head(Beerdata)
nrow(Beerdata)
```

```{r Brewery Data}
# This reads in the Beers data from select folder file Breweries.csv.
Breweriesdata <- read.csv("https://raw.githubusercontent.com/VenkataVanga/MSDS-6306/main/Breweries.csv",header = T,sep = ",",na.strings = "NA",
                          fill = TRUE);
head(Breweriesdata)
nrow(Breweriesdata)
```

@@

#Generating geographic plots from breweries data
```{r mapdata Breweries}
#makes a data frame with State name and abbreviation.
lookup = data.frame(abb = state.abb, State = state.name)
dc <- c("DC", "District of Columbia")
lookup <- rbind(lookup,dc)
# Change Column Name State to abb (abbreviation)
colnames(Breweriesdata)[4] = "abb"

@@

#Code below adds scaled legend to the Breweries count by state
cplot <- Nplot + theme(legend.background = element_rect(fill="gray90", size=10, linetype="dotted"),legend.key.height = unit(4,"lines"))

#final plot with state names and Breweries count by state
cplot

```
#Merging beer data and breweries data
```{r merging beer and breweries data}

@@

>>>>

---
title: "MSDS 6306: Doing Data Science - Case Study 01 - Beers/Breweries Dataset"
author: "Renfeng Wang & Venkata 'MG' Vanga"
date: "9/30/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

@@

#R libraries
```{r libraries}
library(knitr)
library(yaml)
library(dplyr)
library(ggplot2)
library(tidyr)
require(graphics)
library(tidyverse)
library(magrittr)
library(GGally)
library(ggthemes)
library(plotly)
library(maps)
library(stringr)
library(stringi)
library(mapproj)
library(RCurl)
library(readr)
library(rio)
library(naniar)
library(scales)
library(grid)
```

#Beer and Breweries data input from CSV files
```{r Beer Data}
# This reads in the Beers data from select folder file Beers.csv.
Beerdata <- read.csv("https://raw.githubusercontent.com/VenkataVanga/MSDS-6306/main/Beers.csv",
                     header = T,sep = ",",na.strings = "NA",fill = TRUE)
head(Beerdata)
```
#Number of rows in Beer data
```{r number of beer data}
nrow(Beerdata)
```

```{r Brewery Data}
# This reads in the Beers data from select folder file Breweries.csv.
Breweriesdata <- read.csv("https://raw.githubusercontent.com/VenkataVanga/MSDS-6306/main/Breweries.csv",
                          header = T,sep = ",",na.strings = "NA", fill = TRUE)
head(Breweriesdata)
```

#Number of rows in Breweries data
```{r number of breweries data}
nrow(Breweriesdata)
```

@@

#Generating geographic plots from breweries data (following chunk generate the final plot)
```{r mapdata Breweries}
#makes a data frame with State name and abbreviation.
lookup = data.frame(abb = state.abb, State = state.name)
dc <- c("DC", "District of Columbia")
lookup <- rbind(lookup,dc)
# Change Column Name State to abb (abbreviation)
colnames(Breweriesdata)[4] = "abb"

@@

#Code below adds scaled legend to the Breweries count by state
cplot <- Nplot + theme(legend.background = element_rect(fill="gray90", 
                                                        size=10, linetype="dotted"),
                       legend.key.height = unit(4,"lines")) + 
  theme(plot.title = element_text(size=14, face= "bold", colour= "black"))
```

#Final plot with state names and Breweries count by state
```{r map breweries, fig.width=35, fig.height=20, dpi = 240}
cplot  #plot best viewed in scaled full screen zoom 1920px x 1080px (21" or higher dim screen)
```
The final plot shows that 'Colorado' has the highest (47)number of breweries followed by 'California (39)', 'Michigan(32)' and 'Oregon(29)' being the next three successors.

#Merging beer data and breweries data
```{r merging beer and breweries data}
colnames(Beerdata)[5] = "Brew_ID" #Changing columun name to combine both beer data and breweries data
Totaldata <- merge(Beerdata,Breweriesdata, by="Brew_ID", all.x = T)
colnames(Totaldata)[2] = "Beer_Name" #Changing the column name to Beer name
colnames(Totaldata)[8] = "Brewery_Name" # Changing the column name to Brewery name
colnames(Totaldata)[10] = "State" # Changing the column name back to State
```

#First 6 observations after merging beer data and breweries data
```{r First 6 observations after merging}
head(Totaldata,6)
```

#Last 6 observations after merging beer data and breweries data
```{r Last 6 observations after merging}
tail(Totaldata,6)
```

#Missing Values in ABV/IBU columns
```{r Missing Values in ABV/IBU columns}
table(is.na(Totaldata$ABV))# True value represents NA values in the column ABV in original data
table(is.na(Totaldata$IBU))# True value represents NA values in the column IBU in original data
gg_miss_var(Totaldata)+ ggtitle("Missing data values by Category")
```

Graph and the values in table above provide the missing data values in ABV (65) and IBU (1005) columns from originally considered beer data.

#Median for ABV and IBU content for each state (final plot given in below chunk)
```{r Median ABV and IBU content for each state}
#Evaluating median of ABV and IBU (all NA values removed from data)
Median_ABV_IBU <- Totaldata %>% arrange(State) %>% group_by(State) %>% 
  summarize(Median_ABV = median(ABV, na.rm = TRUE), Median_IBU = median(IBU,na.rm = TRUE))
Median_ABV_IBU <- data.frame(Median_ABV_IBU)

#Bar plot for median of ABV content for each state (This plot is added to IBU med plot below)
ABV_Medplot <- Median_ABV_IBU %>% ggplot() + 
  geom_bar(aes(x = State, y = Median_ABV, fill = State),group = Median_ABV_IBU$State, 
           width = 0.3, stat = 'identity', show.legend = FALSE)

#Bar plot for median of IBU content for each state (added to median of ABV plot)
ABV_IBU_Medplot <- ABV_Medplot + geom_bar(aes(x = as.numeric(factor(State)) + 0.4, 
                                              y = Median_IBU/1000),stat = 'identity',width = 0.3)
#Median IBU scaled down by 100 in order to set same y axis scale for ABV and IBU.
#Adding secondary y axis to the plot
ABV_IBU_Medplot1 <- ABV_IBU_Medplot + 
  scale_y_continuous(limits= c(0,0.072), labels = percent, name = "Alcohol By Volume (ABV) [%]", 
                      sec.axis = sec_axis(~.*1000,name = "International Bitterness Unit (IBU)")) +
  ggtitle('Median ABV and IBU Content by State') + theme(
# AXIS LABLES APPEARANCE
plot.title = element_text(size=14, face= "bold", colour= "black" ),
axis.title.x = element_text(size=12, face="bold", colour = "black"),    
axis.title.y = element_text(size=12, face="bold", colour = "black"),    
axis.text.x = element_text(size=12, face="bold", colour = "black"), 
axis.text.y = element_text(size=12, face="bold", colour = "black"),
strip.text.x = element_text(size = 10, face="bold", colour = "black" ),
strip.text.y = element_text(size = 10, face="bold", colour = "black"),
)

#Adding text to median values for ABV and IBU
ABV_IBU_Medplot2 <- ABV_IBU_Medplot1 + geom_text(aes(State, Median_ABV+0.0045, 
                                                     label = paste("ABV =",percent(Median_ABV, 
                                                                                   accuracy = 0.001)),
                                                     fill = NULL),
                                                 data = Median_ABV_IBU, angle = 90) +
  geom_text(aes(as.numeric(factor(State)) + 0.4, (Median_IBU/1000)+0.0032, 
                label = paste("IBU =",Median_IBU), fill = NULL), data = Median_ABV_IBU, angle = 90)

@@

# Median plot for ABV and IBU by state
```{r Final Median Plot for ABV and IBU by state,fig.width=30, fig.height=15, dpi = 240}
ABV_IBU_Medplot2 
#plot best viewed in scaled full screen zoom 1920px x 1080px (21" or higher dim screen)
```

From the median ABV and median IBU plot above it can be observed that 'District of Columbia' and 'Kentucky' have the highest median ABV values shown in percentage. 
'Maine' and 'West Virginia' have the highest median IBU values.

# Maximum ABV and IBU by State
```{r maximum ABV and IBU State}
#collecting max data
Max_ABV <- Totaldata %>% arrange(State) %>% group_by(State) %>% 
  summarize(Max_ABV = max(ABV, na.rm=TRUE))
Max_IBU <- Totaldata %>% arrange(State) %>% group_by(State) %>% 
  summarize(Max_IBU = max(IBU, na.rm=TRUE))

Max_ABV$Cat <- "Max_ABV"
Max_IBU$Cat <- "Max_IBU"

#Bar plot for max ABV content for each state
Maxplot_ABV_IBU <- Max_ABV %>% ggplot(mapping = aes(x,y)) + 
  geom_bar(aes(x = State, y = Max_ABV, fill = State), group = Max_ABV$State, 
                                                   stat = 'identity', show.legend = FALSE) +
  ggtitle('Max ABV Content by State') + geom_text(aes(State, Max_ABV+0.0085, 
                                                      label = percent(Max_ABV, accuracy = 0.01),
                                                      fill = NULL), data = Max_ABV, angle = 90) + 
  geom_bar(data=Max_IBU, aes(x = State, y = Max_IBU, fill = State), stat = 'identity', 
           show.legend = FALSE) +
  geom_text(aes(State, Max_IBU+5, label = Max_IBU, fill = NULL), data = Max_IBU) +
  ggtitle('Max ABV and Max IBU Content by State') + facet_grid(Cat~., scale = "free_y",
                                                              labeller = label_parsed, switch = "y") + labs(x="State") + 
  theme(axis.title.y = element_blank(), strip.placement = "outside")

# Maximum ABV and IBU vaule by state
```{r Final Max ABV and IBU by state,fig.width=30, fig.height=20, dpi = 240}
Maxplot_ABV_IBU 
#plot best viewed in scaled full screen zoom 1920px x 1080px (21" or higher dim screen)
```

From the Max ABV and IBU plot it is observed that 'Colorado' has the max ABV value of 12.8% and 'Oregon' has the max IBU value of 138.

# ABV statistical attributes
```{r}
Stats <- Totaldata %>% summarize(Mean = mean(ABV, na.rm=TRUE),
                        Median = median(ABV,na.rm=T), Max = max(ABV,na.rm=T), Min = min(ABV,na.rm=T),
                        SD = sd(ABV,na.rm=T), N = n())
#Histogram and Density Plot
His_Den <- Totaldata %>% filter(!is.na(ABV)) %>% ggplot(aes(x=ABV)) +
  geom_histogram(aes(y=..density..),colour='red',fill='blue', binwidth = 0.0035) +
  geom_density(alpha=.4, fill='#FFFF00') + 
  ggtitle('ABV Statistical Attributes - Histogram, Density and Box Plots') + 
  scale_x_continuous(expand = c(0,0), limit = c(0,0.14)) + labs(y="Density / Count")

#Box plot for ABV data
Box <- Totaldata %>% filter(!is.na(ABV)) %>% ggplot(aes(x=ABV)) + 
  geom_boxplot(col='black',fill='#FF6666') + scale_x_continuous(expand = c(0,0), limit = c(0,0.14)) + scale_y_continuous(expand = c(0,0), limit = c(-0.4,0.4))

#Histogram density plot and Box plot on same scale
grid.draw(rbind(ggplotGrob(His_Den),
                ggplotGrob(Box),
                size = "first"))
Stats
```
All statically significant attributes for the ABV values are shown above
The distribution of ABV is slightly right skewed. ABV of beers around 5% has the highest count. 
There are total 2410 non-missing ABV values in this data set. 
The maximum ABV is 12.8%, the minimum ABV is .1%, the median ABV is 5.6%.
The mean ABV is 5.98%, median 5.6% and standard deviation of ABV is 1.35%.
